---
title: "community_phenology"
author: "Jannet Vu"
date: "10/26/2022"
output: html_document
---
Manuscript title: Temperature and rainfall seasonality drives tropical community phenology in southeastern Madagascar

**Goal is to characterize the interspecific and intraspecific synchrony of reproductive phenological cycle (flower and fruit) of plants in Kianjavato **\

#### Set up working environment
First set up working environment and load in relevant files
```{r, setup, echo=F, message=F, warning=F}
#set up working directory
setwd("C:/community_phenology_mada")
```

```{r, echo=T, message=F, warning=F}
#load in packages
library(dplyr)
library(lubridate)
library(ggplot2)
library(reshape)
library(doParallel)
library(gridExtra)
library(cowplot)

# detect cores for parallel computing
numcores <- detectCores() -1
registerDoParallel(numcores) # register cores

# load appropriate files
# load in species level phenology data 
# data for characterizing intraspecific synchrony
kphen_fr_sync <- read.csv('kian_phen_popindividual_fruit_sync.csv', header =T, stringsAsFactors = F)[,-1]
kphen_fr_sync$Date <- as.Date(kphen_fr_sync$Date, tz = 'Africa/Nairobi')

kphen_fl_sync <- read.csv('kian_phen_popindividual_flower_sync.csv', header = T, stringsAsFactors = F)[,-1]
kphen_fl_sync$Date <- as.Date(kphen_fl_sync$Date, tz = 'Africa/Nairobi')
```
columns:\
species - vernacular name\
year - year\
month - month\
dataset - forest identifier (sssw - sangasanga, vvvw - vatovavy)\
Tag - individual identifier\
Plot- survey plot\
FuFr - fruit phenology (0-4)\
Fl - flower phenology (0-4)\
fruit - absence/presence of fruit\
flower - absence/presence of flower \
Date - date
monthname - categorical month name
Latin - scientific name(s)\
multispecies - indicates the taxonomic ambiguity, FALSE = single species, TRUE = multiple species, AMB - single species, but uncertain which species
stage - flower or fruit stage identifier

```{r, echo = T, warning = F, message = F}
head(kphen_fr_sync) # output to get sense of data
```

```{r, echo = T, warning = F, message = F}
# data for characterizing species phenology and interspecific synchrony
kphen_sp_fr <- read.csv('kian_phen_popspecies_fruit_sync.csv', header= T, stringsAsFactors = F)[,-1] 
kphen_sp_fr$Date <- as.Date(kphen_sp_fr$Date, tz = 'Africa/Nairobi')

kphen_sp_fl <- read.csv('kian_phen_popspecies_flower_sync.csv', header= T, stringsAsFactors = F)[,-1]
kphen_sp_fl$Date <- as.Date(kphen_sp_fl$Date, tz = 'Africa/Nairobi')

head(kphen_sp_fr) # visualize to get sense of data
```

columns:\
year - year\
month - month\
monthname - categorical month name\
Date - date \
species - vernacular name \
dataset - forest identifier (sssw - sangasanga, vvvw - vatovavy)\
Latin - scientific name\
multispecies - indicates the taxonomic ambiguity, FALSE = single species, TRUE = multiple species, AMB - single species, but 
n - number of individuals sampled
Fl or FuFr - flower or fruit phenology (0-4)\
flower or fruit - absence/presence of flower or fruit\
flower_prop or fruit_prop - proportion of individuals flowering or fruiting

```{r, echo = T, warning = F, message = F}
# load in species level phenology data 
kphen_sp_fr <- read.csv('kian_phen_popspecies_fruit_sync.csv', header= T, stringsAsFactors = F)[,-1] 
kphen_sp_fr$Date <- as.Date(kphen_sp_fr$Date, tz = 'Africa/Nairobi')

kphen_sp_fl <- read.csv('kian_phen_popspecies_flower_sync.csv', header= T, stringsAsFactors = F)[,-1]
kphen_sp_fl$Date <- as.Date(kphen_sp_fl$Date, tz = 'Africa/Nairobi')

head(kphen_sp_fr) # output to get sense of data
```
columns:\
year - year\
month - month\
monthname - categorical month name\
Date - date \
species - vernacular name \
dataset - forest identifier (sssw - sangasanga, vvvw - vatovavy)\
Latin - scientific name\
multispecies - indicates the taxonomic ambiguity, FALSE = single species, TRUE = multiple species, AMB - single species, but uncertain which species
n - number of individuals sampled
Fl or FuFr - mean flower or fruit phenology (0-4)\
flower or fruit - absence/presence of flower or fruit\
flower_prop or fruit_prop - proportion of individuals flowering or fruiting

#### Examine intraspecific and interspecific flowering and fruiting synchrony 
Calculate community synchrony using Loreau's synchrony function\
Loreau et al. 2008. Species Synchrony and Its Drivers: Neutral and Nonneutral Community Dynamics in Fluctuating Environments. The American Nauralist.\
\
Functions for community synchrony calculations
```{r, echo = T, message = F, warning=F}
# community.sync function from the "synchorny" package
community.sync.aux <- function (data) {
  # data = dataframe phenology time series, rows are time points, columns are unique individuals or species 
  species.sd=apply(data, MARGIN=2, FUN=sd) # within species or individual trait SD 
  community.var=var(rowSums(data)) # variance in trait across community or population
  return(community.var/sum(species.sd, na.rm=TRUE)^2) # return Loreau's synchrony metric 
}

# permutation analysis lag function
# perumation analysis will be used to examine whether degree of synchrony deviates from random expectation
perm_lagfun <- function(df){
  # df - dataframe phenology time series, rows are time points, columns are unique individuals or species 
  shift_id <- rep(1:nrow(df), 2) # get time points and append with itself. This will be used to shift the time series
  # we appended with itself so that shifts that occur later can loop back to the beginning 
  perm_factor <- sample(nrow(df), ncol(df), replace = T) # sample time window shift for each column (individual or species)
  perm_df <- sapply(1:ncol(df), function(x) { #  for each column (individual or species)
          perm_end <- perm_factor[x]+nrow(df)-1 # pull the indices according to the randomly sampled time shift
          val_shifts <- shift_id[perm_factor[x]:perm_end] # pull the values based on the time shifted indices 
          return(df[val_shifts, x])
        })
  sync_val <- community.sync.aux(perm_df) # calculate the synchrony function on permuated data
  return(sync_val)
}

# pvalue calculation
perm_pvalue <- function(ref, perms,sp, dataset, st){
  # ref = comm synchrony for observed 
  # perms = comm synchrony based on permutation shifts
  # sp = species
  # dataset = community
  # st = stage (flower or fruit)
  perms <- c(perms,ref) # append reference synchrony to synchrony calcalated from permutations
  perms_shift <- perms - mean(perms) # subtract from mean to get test statistic
  ref_shift <- ref - mean(perms) # subtract from mean to get test statistic
  pval <- sum(abs(perms_shift) >= abs(ref_shift))/length(perms_shift) # get proportion of values that are more extreme than the absolute value of the observed metric, since a two tailed test is performed
  # format output info into dataframe
  pval_df <- data.frame(species = sp, 
                        community = dataset, 
                        stage = st, 
                        comm_sync = ref, 
                        random_sync_mean = mean(perms), 
                        random_sync_sd = sd(perms), 
                        sync_diff = ref_shift, 
                        sync_magnitude = ref/mean(perms), 
                        pvalue = pval) # output info
  return(pval_df)
}

# wrapper function for synchrony calculation and permutation analysis
comm_sync_wrapper <- function(df, sp, dataset, st, nperms = 9999){
  # df - dataframe phenology time series, rows are time points, columns are unique individuals or species 
  # sp = species
  # dataset = community
  # st = stage (flower or fruit)
  # nperms = number of permutation iterations to run
  ref <- community.sync.aux(df) # calculate the reference community synchrony
  perms <- c() # create empty vector for synchrony values from permuted data
  for(i in 1:nperms){
    perms[i] <- perm_lagfun(df) # get permuted synchrony values 
  }
  output <- perm_pvalue(ref, perms, sp, dataset, st) # calculate p-value
  return(output)
}
```

Calculate intraspecific synchrony
```{r, echo = T, message = F, warning=F}
# create empty dataframe to populate with synchrony info
sync_df <- data.frame(matrix(ncol = 9, nrow = 0)) # create empty dataframe
colnames(sync_df) <- c('species','community','stage',
                       'comm_sync','random_sync_mean',
                       'random_sync_sd', 'sync_diff',
                       'sync_magnitude','pval') # generate column names

# make sure in chronological order
kphen_fr_sync <- arrange(kphen_fr_sync, Tag, Date)
kphen_fl_sync <- arrange(kphen_fl_sync, Tag, Date)

# Transform each species into a separate list 
data_list <- lapply(split(kphen_fr_sync, ~Tag),function(x) as.list(x)) # transform dataframe into list
#make dataframe of timeseries
data_ls <-plyr::llply(data_list,.fun=function(x) ts(x$FuFr,start=c(x$year[1],x$month[1]),freq=12, class = 'ts')) # Output phenology in list form
id_df <- distinct(kphen_fr_sync[,c('Tag','species','dataset')]) # get tag-species association for subsetting

# for each species and each dataset stage
for (sp in unique(kphen_fr_sync$species)){ # for each species
  indices_ss <- which(id_df$species == sp & id_df$dataset =='sssw') # get indices of the tags that pertain to the species for Sangasanga
  indices_vv <- which(id_df$species == sp & id_df$dataset =='vvsw') # get indices of the tags that pertain to the species for Vatovavy
  if(length(indices_ss) > 2){ # if there are at least 3 individuals 
    sp_wide <- Reduce(ts.intersect, data_ls[indices_ss]) # get the time series of that species in a wide-format table (rows: times, columns: individuals )
    sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, sp, 'sssw','fruit') # add to synchrony table
  }
  if(length(indices_vv) > 2){ # if there are at least 3 individuals 
    sp_wide <- Reduce(ts.intersect, data_ls[indices_vv]) # get the timeseries of that species in a wide-format table (rows: times, columns: individuals )
    sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, sp, 'vvsw','fruit') # add to synchrony table
  }
}
```

Repeat for intraspecific flowering synchrony 
```{r, echo = T, message = F, warning=F}
data_list <- lapply(split(kphen_fl_sync, ~Tag),function(x) as.list(x)) # transform dataframe into list
#make dataframe of timeseries "data_df"
data_ls <-plyr::llply(data_list,.fun=function(x) ts(x$Fl,start=c(x$year[1],x$month[1]),freq=12, class = 'ts')) # Output phenology in list form
id_df <- distinct(kphen_fl_sync[,c('Tag','species','dataset')])

for (sp in unique(kphen_fl_sync$species)){ # for each species
  indices_ss <- which(id_df$species == sp & id_df$dataset =='sssw') # get indices of the tags that pertain to the species for Sangasanga
  indices_vv <- which(id_df$species == sp & id_df$dataset =='vvsw') # get indices of the tags that pertain to the species for Vatovavy
  if(length(indices_ss) > 2){ # if there are at least 3 individuals 
    sp_wide <- Reduce(ts.intersect, data_ls[indices_ss]) # get the timeseries of that species in a wide-format table (rows: times, columns: individuals )
    sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, sp, 'sssw','flower') # add to synchrony table
  }
  if(length(indices_vv) > 2){ # if there are at least 3 individuals 
    sp_wide <- Reduce(ts.intersect, data_ls[indices_vv]) # get the timeseries of that species in a wide-format table (rows: times, columns: individuals )
    sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, sp, 'vvsw','flower') # add to synchrony table
  }
}

# organize data by stage and community
sync_df <- arrange(sync_df, stage, community)
write.csv(sync_df, 'phen_species_sync_output.csv')
```

Calculate for interspecific synchrony
```{r, echo = T, message = F, warning=F}
# make sure in chronological order
kphen_sp_fl <- arrange(kphen_sp_fl, species, Date)
kphen_sp_fr <- arrange(kphen_sp_fr, species, Date)

# Analyze interspecific flowering synchrony for Vatovavy using productivity
data_list <- lapply(split(kphen_sp_fl %>% filter(dataset =='vvsw'), ~species),
                    function(x) as.list(x)) # transform dataframe into list
data_ls <-plyr::llply(data_list,.fun=function(x) ts(x$Fl,start=c(x$year[1],x$month[1]),freq=12, class = 'ts')) # Output phenology in list form
sp_wide <- Reduce(ts.intersect, data_ls) # convert to wide format
sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, 'community', 'vvsw','flower') # add to synchrony table

# Analyze interspecific flowering synchrony for Sangasanga using productivity
data_list <- lapply(split(kphen_sp_fl %>% filter(dataset =='sssw'), ~species),
                    function(x) as.list(x)) # transform dataframe into list
data_ls <-plyr::llply(data_list,.fun=function(x) ts(x$Fl,start=c(x$year[1],x$month[1]),freq=12, class = 'ts')) # Output phenology in list form
sp_wide <- Reduce(ts.intersect, data_ls) # convert to wide format
sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, 'community', 'sssw','flower') # add to synchrony table


# Analyze interspecific fruiting synchrony for Vatovavy using productivity
data_list <- lapply(split(kphen_sp_fr %>% filter(dataset =='vvsw'), ~species),
                    function(x) as.list(x)) # transform dataframe into list
data_ls <-plyr::llply(data_list,.fun=function(x) ts(x$FuFr,start=c(x$year[1],x$month[1]),freq=12, class = 'ts')) # Output phenology in list form
sp_wide <- Reduce(ts.intersect, data_ls) # convert to wide format
sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, 'community', 'vvsw','fruit') # add to synchrony table

# Analyze interspecific fruiting synchrony for Sangasanga using productivity
data_list <- lapply(split(kphen_sp_fr %>% filter(dataset =='sssw'), ~species),
                    function(x) as.list(x)) # transform dataframe into list
data_ls <-plyr::llply(data_list,.fun=function(x) ts(x$FuFr,start=c(x$year[1],x$month[1]),freq=12, class = 'ts')) # Output phenology in list form
sp_wide <- Reduce(ts.intersect, data_ls) # convert to wide format
sync_df[nrow(sync_df)+1,] <- comm_sync_wrapper(sp_wide, 'community', 'sssw','fruit') # add to synchrony table

write.csv(sync_df, 'phen_species_sync_output.csv')

sync_df
```
