---
title: "community_phenology"
author: "Jannet Vu"
date: "10/26/2022"
output: html_document
---
Manuscript title: Temperature and rainfall seasonality drives tropical community phenology in southeastern Madagascar

**Goal is to characterize the reproductive phenological cycle (flower and fruit) of plants in Kianjavato at the individual and species level at the monthly scale**\
1) identify the dominant cycle flowering and fruiting frequency (continuous, sub-annual, annual, supra-annual)\
2) calculate the flowering and fruiting cycle length, duration and interval \
3) assess the regularity of flowering and fruiting 

#### Set up working environment
First set up working environment and load in relevant files
```{r, setup, echo=T, message=FALSE, warning=FALSE}
# set working directory for all chunks
setwd("C:/Users/Jannet/Documents/Dissertation/data")
knitr::opts_knit$set(root.dir = "C:/Users/Jannet/Documents/Dissertation/data") 
```


```{r, setup, echo=T, message=FALSE, warning=FALSE}
#load in packages
library(dplyr)
library(lubridate)
library(ggplot2)
library(reshape)
library(doParallel)
library(gridExtra)
library(cowplot)
library(chisq.posthoc.test)

# set up working directory 
setwd("C:/data")

# set up directory to save figures
figpath <- "C:/figures"

# detect cores for parallel computing
numcores <- detectCores() -1
registerDoParallel(numcores) # register cores

# list of month labels 
mthname <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")

# load appropriate files
# data for plant vernacular and Latin names
plant_names <- read.csv('phen_plantnames.csv', header= T, stringsAsFactors = F) # local plant and Latin names

# data for characterizing individual fruiting cycles
kphen_fr <- read.csv('kian_phen_individual_fruit.csv', header =T, stringsAsFactors = F)[,-1]
kphen_fr$Date <- as.Date(kphen_fr$Date, tz = 'Africa/Nairobi')

# data for characterizing individual flowering cycles
kphen_fl <- read.csv('kian_phen_individual_flower.csv', header = T, stringsAsFactors = F)[,-1]
kphen_fl$Date <- as.Date(kphen_fl$Date, tz = 'Africa/Nairobi')
```
columns:\
species - vernacular name\
year - year\
month - month\
dataset - forest identifier (sssw - sangasanga, vvvw - vatovavy)\
Tag - individual identifier\
Plot- survey plot\
Fu - unripe fruit phenology (0-4)\
Fr - ripe fruit phenology (0-4)\
FuFr - fruit phenology (0-4)\
Fl - flower phenology (0-4)\
fruit - absence/presence of fruit\
flower - absence/presence of flower \
Date - date
monthname - categorical month name
Latin - scientific name(s)\
multispecies - indicates the taxonomic ambiguity, FALSE = single species, TRUE = multiple species, AMB - single species, but uncertain which species

```{r, echo = T, warning = F, message = F}
head(kphen_fr) # visualize to get sense of data
```

**What is the starting sample size for the individual analyses?**\
Data on fruiting phenology was collected on: `r length(unique(kphen_fr$Tag))` individuals from `r length(unique(kphen_fr$Species))` taxa.\
Date on flowering phenology was collected on: `r length(unique(kphen_fl$Tag))` individuals from `r length(unique(kphen_fl$Species))` species.\

```{r, echo = T, warning = F, message = F}
# load in species level phenology data 
kphen_sp_fr <- read.csv('kian_phen_popspecies_fruit_sync.csv', header= T, stringsAsFactors = F)[,-1] 
kphen_sp_fr$Date <- as.Date(kphen_sp_fr$Date, tz = 'Africa/Nairobi')

kphen_sp_fl <- read.csv('kian_phen_popspecies_flower_sync.csv', header= T, stringsAsFactors = F)[,-1]
kphen_sp_fl$Date <- as.Date(kphen_sp_fl$Date, tz = 'Africa/Nairobi')
```
columns:\
year - year\
month - month\
monthname - categorical month name\
Date - date \
species - vernacular name \
dataset - forest identifier (sssw - sangasanga, vvvw - vatovavy)\
Latin - scientific name\
multispecies - indicates the taxonomic ambiguity, FALSE = single species, TRUE = multiple species, AMB - single species, but uncertain which species
n - number of individuals sampled
Fl or FuFr - flower or fruit phenology (0-4)\
flower or fruit - absence/presence of flower or fruit\
flower_prop or fruit_prop - proportion of individuals flowering or fruiting

```{r, echo = T, warning = F, message = F}
head(kphen_sp_fr) # visualize to get sense of data
```

**What is the starting sample size for the species analyses?**\
The fruiting phenology of at least 3 individuals was recorded for `r length(unique(kphen_sp_fr$Species))` species.\
The flowering phenology of at least 3 individuals was recorded for `r length(unique(kphen_sp_fl$Species))` species.

Merge flowering fruiting individual and species tables
```{r, echo = T, warning = F, message = F}
# assign phenology attributes before merging tables 
kphen_fl$stage <- 'flower'
kphen_fr$stage <- 'fruit'
kphen_sp_fl$stage <- 'flower'
kphen_sp_fr$stage <- 'fruit'
# assign tag as the forest species was sampled since species doesn't have tag number
kphen_sp_fl$Tag <- kphen_sp_fl$dataset
kphen_sp_fr$Tag <- kphen_sp_fr$dataset

# assign flower prop and fruit prop for individual as NA before merging, since it doesn't exist for individuals
kphen_fl$flower_prop <- NA
kphen_fr$fruit_prop <- NA

# merge flower individual and species tables first 
# list out relevant columns to merge by
relecols <- c('species','year','month','dataset','Tag','Fl','flower',
              'flower_prop','Date','monthname','Latin','multispecies','stage')
fl_merge <- rbind(kphen_fl[,relecols], kphen_sp_fl[,relecols]) # row bind flowering species and individual data 
colnames(fl_merge)[6:8] <- c('productivity','presence','proportion') # rename columns

# merge fruit individual and species tables next 
relecols <- c('species','year','month','dataset','Tag','FuFr','fruit',
              'fruit_prop','Date','monthname','Latin','multispecies','stage')
fr_merge <- rbind(kphen_fr[,relecols], kphen_sp_fr[,relecols]) # row bind fruiting species and individual data 
colnames(fr_merge)[6:8] <- c('productivity','presence','proportion') # rename columns

# merge fruit and flower individual and species data together 
phen_merged <- rbind(fl_merge, fr_merge)

# generate table of of the cycle characterization for each individual
# assign unique identifier
phen_merged$ID <- sapply(1:nrow(phen_merged), 
                         function(x) paste0(phen_merged[x,c('species','Tag','stage')], collapse ="_"))

phen_merged <- arrange(phen_merged, ID, Date) # order the table 
```

Visualize sample species phenology 
```{r, echo = T, warning = F, message = F}
# get a random species
subsp <- sample(phen_merged$species, 1) 
# subset out sample species flowering data 
subdf <- phen_merged[which(phen_merged$species == subsp),]
subdf <- arrange(subdf, Date) # arrange chronologically

# plot time series of sample species flower productivity  
fl_plot <- ggplot(subdf %>% filter(stage == 'flower'),aes(x=Date,y=productivity))+
                geom_line(aes(color = dataset), size = 1)+
                ylab("productivity") +
                xlab("") +
                theme(legend.position="bottom",
                      panel.background = element_rect(fill = 'white',colour = 'black'),
                      panel.grid.major.x =  element_line(colour = "grey80"),
                      panel.grid.minor.x = element_line(colour = "grey80")) +
                facet_wrap(~Tag,ncol=1, strip.position = 'right', scales = 'free_y')+
                scale_color_brewer(palette = 'Accent') +
                ggtitle(paste(subsp, 'flower'))

# plot time series of sample species fruit productivity  
fr_plot <- ggplot(subdf %>% filter(stage == 'fruit'),aes(x=Date,y=productivity))+
                geom_line(aes(color = dataset), size = 1)+
                ylab("presence")+
                xlab("")+
                theme(legend.position="bottom") + 
                theme(panel.background = element_rect(fill = 'white',colour = 'black'),
                panel.grid.major.x =  element_line(colour = "grey80"),
                panel.grid.minor.x = element_line(colour = "grey80")) +
                facet_wrap(~Tag,ncol=1, strip.position = 'right', scales = 'free_y')+
                scale_color_brewer(palette = 'Accent') +
                ggtitle(paste(subsp, 'fruit'))

# plot sample flower and fruit productivity side by side
grid.arrange(fl_plot,fr_plot,
             nrow = 1, 
             ncol = 2)
```

Function to calculate phenological cycle duration, interval and length for each individual and species
```{r,echo = T,warning=F,message = F}
cycle_quant <- function(cycle_component, ID){
  # cycle_component: run length encoder that calculates number 
  # consecutive flowering/fruiting months (duration) and nonflowering/nonfruiting months (intervals)
  # ID: unique individual or species identifier
  # get number of intervals and durations
  # extract out whether the cycle length pertains to a flowering interval (0) or duration 1)
  cycle_values <- cycle_component$values
  # extract out the number of months for each flowering duration/interval
  cycle_lengths <- cycle_component$lengths
  
  # get number of cycle components
  num_cycles <- length(cycle_component$lengths) 
  # get number of flowering/fruiting durations (as indicated by a value of 1)
  on_months <- sum(cycle_component$lengths[which(cycle_component$values == 1)]) 
  # get number of flowering/fruiting intervals (as indicated by a value of 0)
  off_months <- sum(cycle_component$lengths[which(cycle_component$values == 0)])
  # get number of episodes (an episode is a full cycle, 
  # which includes a duration of flowering/fruiting and nonflowering/nonfruiting
  episodes <- floor((on_months + off_months)/2)
  episodes <- sum(cycle_component$values,na.rm =T) 
  
  # summarize cycle characteristics
  if (length(cycle_lengths) < 2) { # if there is less than 2 cycles  
    # don't have full cycle, set to NA
    cycle_df <- data.frame(ID = ID, 
                           episodes = episodes,
                           cycle_num = 0, 
                           total_duration = on_months, 
                           total_interval = off_months,
                           cycle_duration = on_months, 
                           cycle_interval = off_months,
                           start = NA) # start "on" indicates that a record 
                                       # started with an active flowering or fruiting period
                                       # and "off indicates that a record started with a 
                                       # nonflowering or nonfruiting period
    print('cycle length too short to quantify')
  }
  
  else { # if there is more than 2 cycles
    if (cycle_values[1] == 0) { # if the first entry is a nonfruiting or nonflowering period
      # drop the first records to ensure the sequence starts 
      # with a active fruiting or flowering period
      cycle_lengths <- cycle_lengths[2:length(cycle_lengths)] 
      cycle_values <- cycle_values[2:length(cycle_values)]
      print('drop first nonfruiting, nonflowering record')
      
      if (length(cycle_lengths) < 2) { # if the number of periods is fewer than two 
        # don't have full cycle, set to NA
        cycle_df <- data.frame(ID = ID, 
                               episodes = episodes, 
                               cycle_num = 0, 
                               total_duration = on_months, 
                               total_interval = off_months,
                               cycle_duration = on_months, 
                               cycle_interval = off_months,
                               start = 'off')
        print('cycle length too short to quantify')
      }
      
      else{
        if (length(cycle_values) %% 2 == 1){ # if the number of periods is odd
          # drop last record to ensure full cycles
          cycle_lengths <- cycle_lengths[1:(length(cycle_lengths)-1)] 
          cycle_values <- cycle_values[1:(length(cycle_values)-1)]
          print('number of periods was odd so dropped last period')
        }
        cycle_number <- sum(cycle_values) # get number of cycles
        cycle_df <- data.frame(ID =ID, 
                               episodes = episodes, 
                               cycle_num = 1:cycle_number, # assign cycle ID 
                               total_duration = on_months, 
                               total_interval = off_months,
                               # get each cycle duration
                               cycle_duration = cycle_lengths[seq(1,length(cycle_lengths),2)], 
                               # get each cycle interval
                               cycle_interval = cycle_lengths[seq(2,length(cycle_lengths),2)],
                               start ='off') 
        print('quantified cycle')
      }
    }
    
    else { # else if the first entry is a fruiting or flowering period 
      if (length(cycle_values) %% 2 == 1) { # if the number of periods is odd
          # drop last record to ensure full cycles
          cycle_lengths <- cycle_lengths[1:(length(cycle_lengths)-1)] 
          print(length(cycle_lengths))
          cycle_values <- cycle_values[1:(length(cycle_values)-1)]
          print('number of periods was odd so dropped last period')
      }
      cycle_number <- sum(cycle_values) # get number of cycles
      cycle_df <- data.frame(ID = ID, 
                             episodes = episodes, 
                             cycle_num = 1:cycle_number, # assign cycle ID 
                             total_duration = on_months, 
                             total_interval = off_months,
                             # get each cycle duration
                             cycle_duration = cycle_lengths[seq(1,length(cycle_lengths),2)], 
                             # get each cycle interval
                             cycle_interval = cycle_lengths[seq(2,length(cycle_lengths),2)],
                             start ='on') 
      print('quantified cycle')
    }
  }
  cycle_df$cycle_length <- cycle_df$cycle_duration + cycle_df$cycle_interval # calculate cycle length
  return(cycle_df) # output the table 
}
```

```{r, echo = T, warning = F, message = F}
# calculate cycle characteristics in parallel for each individual and species in parallel
cycle_df <- foreach(i = unique(phen_merged$ID), .combine = 'rbind') %dopar%{
  # subset the phenology records for an individual/species
  # data should already be chronological 
  subdf <- phen_merged[which(phen_merged$ID == i),] 
  # calculate periods of flowering or fruiting (duration) and nonflowering or nonfruiting (intervals) 
  cycle_component <- rle(subdf$presence) 
  # characterize each flowering/fruiting cycle
  sub_cycle <- cycle_quant(cycle_component, i)
  return(sub_cycle)
}
```

Assign cycle categorization based on the classification scheme in \
Newstrom LE, Frankie GW, Baker HG (1994a) A new classification for plant phenology based on flowering patterns in lowland tropical rain forest trees at La Selva, Costa Rica. Biotropica 26:141–159. https://doi.org/10.2307/2388804\
\
Annual flowering and fruiting cycles occur once per year\
Sub-annual cycles occur multiple times per year\
Supra-annual cycles occur less than once a year\
Continual cycles are patterns of constant flowering or fruiting with no or infrequent, brief interruptions
```{r, echo = T, warning = F, message = F}
# function to categorize cycles
# if cycle length is less than 11 months assign as sub-annual
cycle_df$cycle_category <- ifelse(cycle_df$cycle_length < 11, 'sub-annual', 
                                  # if cycle length is between 11-13 months assign as annual
                                  ifelse(between(cycle_df$cycle_length,11,13),'annual',
                                         # if cycle interval is greater than 2 months, 
                                         # then assign as supra-annual
                                         ifelse(cycle_df$cycle_interval > 2, 'supra-annual', 
                                                # otherwise categorize as continual 
                                                ifelse(cycle_df$cycle_interval <=2, 'continual',NA))))

# for individuals/species that exhibited only one cycle, assign as either supra-annual or continual
# leave the cycle categories of individuals/species with more than one cycle alone
cycle_df$cycle_category <- ifelse(cycle_df$episodes != 1, cycle_df$cycle_category,
                                  # if the cycle duration was shorter than the interval duration
                                  # then asasign as supra-annual, otherwise assign as continual
                                   ifelse(cycle_df$total_duration < cycle_df$total_interval,
                                          'supra-annual','continual')) 

# get rid of leading and tailing sub-annual and annual cycles 
# leading = first cycle for an individual/species
# tailing = last cycle for an individual/species
cycle_df$cycle_category <- ifelse(cycle_df$cycle_num == 1 & # if it is a leading cycle 
                                    # and the cycle is categorized as a sub-annual and annual 
                                    cycle_df$cycle_category %in% c('sub-annual','annual') & 
                                    # and the cycle starts as an active flowering/fruiting period 
                                    # then reassign to NA, otherwise keep as is
                                    cycle_df$start == 'on', NA, cycle_df$cycle_category)

cycle_df$cycle_category <- ifelse(cycle_df$cycle_num == cycle_df$episodes & # if it is a tailing cycle 
                                    # and the cycle is categorized as a sub-annual and annual 
                                    cycle_df$cycle_category %in% c('sub-annual','annual') & 
                                    # and the cycle ends as an active flowering/fruiting period 
                                    # then reassign to NA, otherwise keep as is
                                    cycle_df$start == 'off',NA, cycle_df$cycle_category) 

# add in identifying information for each cycle based on its ID
cycle_df <- merge(cycle_df, distinct(phen_merged[,c('ID','species','Tag','stage','dataset')]), by = 'ID', all.x = T, all.y = F)

# assign the scale of analysis (species or individual)
# Tags = sssw or vvsw indicate species
cycle_df$scale <- ifelse(cycle_df$Tag %in% c('sssw','vvsw'), 'species', 'individual')

# view subsample of cycle data 
head(cycle_df[which(is.na(cycle_df$cycle_category)),])

# get number of cycles in each category for individuals
cycle_df %>% filter(scale != 'species') %>% group_by(cycle_category,dataset) %>% count()

# get number of cycles in each category for species 
cycle_df %>% filter(scale == 'species') %>% group_by(cycle_category,dataset) %>% count()

write.csv(cycle_df, 'phen_cycle_characteristics.csv')
```

Summarize the cycle characteristics to the individual/species
```{r, echo = T, warning = F, message = F}
# create vector of cycle names  
cycle_names <- c('continual','sub-annual','annual','supra-annual')

# get rid of cycles for individuals or species that could not be characterized due to too few cycles
cycle_df <- cycle_df[which(!is.na(cycle_df$cycle_category)),] 

# get summary statistics for each individual and species 
cycle_summary <- cycle_df %>% group_by(ID, species, Tag, dataset, stage,scale, episodes, 
                                       total_duration, total_interval) %>%
                              summarise(n = n(), 
                                        continual = length(which(cycle_category == 'continual')),
                                        sub_annual = length(which(cycle_category == 'sub-annual')),
                                        annual = length(which(cycle_category == 'annual')),
                                        supra_annual = length(which(cycle_category == 'supra-annual')),
                                        length_min = min(cycle_length, na.rm =T), 
                                        length_max = max(cycle_length, na.rm =T), 
                                        length_mean = mean(cycle_length, na.rm =T), 
                                        length_sd = sd(cycle_length,na.rm =T), 
                                        length_median = median(cycle_length,na.rm =T), 
                                        length_mad = mad(cycle_length,na.rm =T),
                                        duration_min = min(cycle_duration, na.rm =T), 
                                        duration_max = max(cycle_duration, na.rm =T), 
                                        duration_mean = mean(cycle_duration, na.rm =T), 
                                        duration_sd = sd(cycle_duration,na.rm =T), 
                                        duration_median = median(cycle_duration,na.rm =T), 
                                        duration_mad = mad(cycle_duration,na.rm =T),
                                        interval_min = min(cycle_interval, na.rm =T), 
                                        interval_max = max(cycle_interval, na.rm =T), 
                                        interval_mean = mean(cycle_interval, na.rm =T), 
                                        interval_sd = sd(cycle_interval,na.rm =T), 
                                        interval_median = median(cycle_interval,na.rm =T), 
                                        interval_mad = mad(cycle_interval,na.rm =T)) %>% 
                              # get dominant cycle frequency and proportion
                              mutate(dom_freq = max(c(continual, sub_annual,annual, supra_annual), na.rm = T), 
                                     dom_prop = max(c(continual, sub_annual,annual, supra_annual), na.rm = T)/n(), 
                                     dom_cycles = paste0(cycle_names[which(c(continual, sub_annual,annual,                                                                                      supra_annual)==dom_freq)], collapse ='/'))

# merge in plant names
cycle_summary <- merge(plant_names, cycle_summary, by.x = 'Species',by.y = 'species', all.x = F, all.y = T)

# flag plants without a dominant cycle 
cycle_summary$cycle_category <- sapply(1:nrow(cycle_summary), 
                                       function(x) grepl('/', cycle_summary$dom_cycle[x]))

# reassign plants without a dominant cycle to "no dominance'
cycle_summary$cycle_category <- ifelse(cycle_summary$cycle_category == 'FALSE', 
                                       cycle_summary$dom_cycle, 'no dominance')

# convert to factor and specify level order
cycle_summary$cycle_category <- factor(cycle_summary$cycle_category, 
                                       levels = c('continual','sub-annual','annual','supra-annual','no dominance'))


# get column names
colnames(cycle_summary)
# reorder table columns
cycle_summary <- cycle_summary[,c(4,1:3,5:8,12,38,37,35:36,13:16,9:11,17:34)]

# get the number of dominant cycles in each category for individuals
cycle_summary %>% filter(scale != 'individual') %>% group_by(stage, dataset, cycle_category) %>% count() 

# get the number of dominant cycles in each category for species
cycle_summary %>% filter(scale == 'species') %>% group_by(stage, dataset, cycle_category) %>% count() 
```

Get a sense of the individual and species sample size before dropping cycles with too few cycles
```{r, echo = T, warning = F, message = F}
# summarize the sample size of the number of cycles completed by species during the study period for flowers
cycle_species_sample <- cycle_summary %>% filter(stage == 'flower' & scale == 'species') %>% 
                                          group_by(n) %>% 
                                          summarise(counts = n()) %>% 
                                          arrange(desc(n))

# get cumulative sample sizes
cycle_species_sample$samples <- cumsum(cycle_species_sample$counts) 

# plot cumulative sample size
fls_plot <- ggplot(cycle_species_sample, aes(x = n, y = samples)) +
                geom_point() + 
                geom_line() + 
                labs(x = 'number of flower cycles', y = 'number of species sampled')

# summarize the sample size of the number of cycles completed by species during the study period for fruit
cycle_species_sample <- cycle_summary %>% filter(stage == 'fruit' & scale == 'species') %>% 
                                          group_by(n) %>% 
                                          summarise(counts = n()) %>% 
                                          arrange(desc(n))

# get cumulative sample sizes
cycle_species_sample$samples <- cumsum(cycle_species_sample$counts) 

frs_plot <- ggplot(cycle_species_sample, aes(x = n, y = samples)) +
                geom_point() + 
                geom_line() + 
                labs(x = 'number of fruit cycles', y = 'number of species sampled') 

# summarize the sample size of the number of cycles completed by individuals during the study period for flowers
cycle_individual_sample <- cycle_summary %>% filter(stage == 'flower' & scale == 'individual') %>% 
                                             group_by(n) %>% 
                                             summarise(counts = n()) %>% 
                                             arrange(desc(n)) 

# cumulative sample size
cycle_individual_sample$samples <- cumsum(cycle_individual_sample$counts) 

# plot cumulative sample size
fli_plot <- ggplot(cycle_individual_sample, aes(x = n, y = samples)) +
                geom_point() + 
                geom_line() + 
                labs(x = 'number of flower cycles', y = 'number of individuals sampled')

# summarize the sample size of the number of cycles completed by individuals during the study period for fruit
cycle_individual_sample <- cycle_summary %>% filter(stage == 'fruit' & scale =='individual') %>% 
                                             group_by(n) %>% 
                                             summarise(counts = n()) %>% 
                                             arrange(desc(n))

# cumulative sample size
cycle_individual_sample$samples <- cumsum(cycle_individual_sample$counts) 

# plot cumulative sample size
fri_plot <- ggplot(cycle_individual_sample, aes(x = n, y = samples)) +
                geom_point() + 
                geom_line() + 
                labs(x = 'number of fruit cycles', y = 'number of individuals sampled')


grid.arrange(fli_plot, fri_plot, fls_plot, frs_plot, nrow =2, ncol = 2)

# # restrict analyses to individuals and species that underwent at least 3 cycles 
# cycle_summary <- cycle_summary[which(cycle_summary$n > 2),]
# write.csv(cycle_summary, 'phen_cycle_character_summary.csv') # save table
# 
# # correspondingly subset the cycle characteristic table with the plants that have at least 3 cycles
# cycle_df <- cycle_df[which(cycle_df$ID %in% cycle_summary$ID),]
```
**The flowering and fruiting phenologies were characterized for how many individuals and species?**\
```{r, echo = T, warning = F, message = F}
cycle_summary %>% group_by(scale, stage,dataset) %>% summarise(n = length(unique(ID)), taxa = length(unique(Species)))
```

Visualize random sample of results
```{r, echo = T, warning = F, message = F}
# draw five random plants
ids <- sample(cycle_summary$ID, 5, replace = F)

# subset out phenology data of the sample 5 plants
subdf <- phen_merged[which(phen_merged$ID %in% ids),]
# subset the cycle summary of these sample plants
subdf <- merge(subdf, cycle_summary[,c('ID','dom_cycles')], by = 'ID', all.x =T, all.y = F)
# chornologically order
subdf <- arrange(subdf, ID, Date)

# plot time series of absence/presene of these sample plants
ggplot(subdf,aes(x=Date,y=presence))+
                geom_line(aes(color = dataset), size = 1)+
                ylab("presence")+
                xlab("")+
                theme(legend.position="bottom") + 
                theme(panel.background = element_rect(fill = 'white',colour = 'black'),
                panel.grid.major.x =  element_line(colour = "grey80"),
                panel.grid.minor.x = element_line(colour = "grey80")) +
                facet_wrap(~paste(ID,dom_cycles, sep = ":"),ncol=1, strip.position = 'top', scales = 'free_y')+
                scale_color_brewer(palette = 'Accent')

```

Visualize distribution of dominant cycles at population level
```{r, echo = T, warning = F, message = F}
# summarize dominant cycles
dompop_summary <- cycle_summary %>% group_by(dataset, stage, scale,cycle_category) %>% 
                                    count(cycle_category,.drop = FALSE)

# tally dominant cycles and get Goodman Estimation of confidence of multinomial proportions
# create empty list to store a table for 
# flowering individual, fruiting individual, flowering species, fruiting species for each forest
mci_list <- list() 

# generate dominant cycle tables for each forest 
for (i in c('sssw','vvsw')){
  flindi <- MultinomCI(dompop_summary$n[which(dompop_summary$dataset == i &
                                              dompop_summary$stage == 'flower' &  
                                              dompop_summary$scale == 'individual')],
                       0.95, 'two.sided', 'goodman') # generate dominany cycle table for flowering individual
  
  frindi <- MultinomCI(dompop_summary$n[which(dompop_summary$dataset == i &
                                              dompop_summary$stage == 'fruit' &  
                                              dompop_summary$scale == 'individual')],
                       0.95, 'two.sided', 'goodman') # generate dominany cycle table for fruiting individual
  
  flsp <- MultinomCI(dompop_summary$n[which(dompop_summary$dataset == i &
                                              dompop_summary$stage == 'flower' &  
                                              dompop_summary$scale == 'species')],
                     0.95, 'two.sided', 'goodman') # generate dominany cycle table for flowering species
  
  frsp <- MultinomCI(dompop_summary$n[which(dompop_summary$dataset == i &
                                              dompop_summary$stage == 'fruit' &  
                                              dompop_summary$scale == 'species')],
                     0.95, 'two.sided', 'goodman') # generate dominany cycle table for fruiting individual
  
  mci_list <- append(mci_list,list(flindi,flsp,frindi,frsp)) # append to list
}

# merge all dominant cycle tables into a single table 
ml_list <- do.call(rbind, mci_list) # merge tables 

# merge multinomial estimates to summary table
dompop_summary <- cbind(as.data.frame(dompop_summary), as.data.frame(ml_list)) 

# print summary table 
print(dompop_summary)

# save table
write.csv(dompop_summary, 'dom_pop_summary_commphen.csv')

# relabel the dataset column with the full forest name for visualization purposes
dompop_summary$dataset <- ifelse(dompop_summary$dataset =='sssw', 'Sangasanga','Vatovavy')

# generate barplot of dominant cycles with confidence intervals for individuals 
indi_plot <- ggplot(dompop_summary %>% filter(scale == 'individual'), 
                  aes(x = dataset, y = est, fill = cycle_category)) +
                  geom_bar(stat = 'identity', position = 'dodge', color = 'black') +
                  geom_errorbar(aes(ymin = lwr.ci, ymax = upr.ci), 
                                position = position_dodge(width = 0.9), colour = 'black', width = 0.2) +
                  theme_bw() +
                  theme(axis.line = element_line(color='black'),
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8),
                        plot.background = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        strip.text.x = element_blank(),
                        legend.position = 'bottom')+
                scale_fill_brewer(palette='Greys') + 
                labs(y = 'proportion', x = 'dominant cycle')+
            facet_wrap(~stage, nrow = 1, ncol = 2) 

# generate barplot of dominant cycles with confidence intervals for species
sp_plot <- ggplot(dompop_summary %>% filter(scale == 'species'), 
                  aes(x = dataset, y = est, fill = cycle_category)) +
            geom_bar(stat = 'identity', position = 'dodge', color = 'black') +
            geom_errorbar(aes(ymin = lwr.ci, ymax = upr.ci), 
                          position = position_dodge(width = 0.9), colour = 'black', width = 0.2) +
            theme_bw() +
            theme(axis.line = element_line(color='black'),
                  text = element_text(size = 8), 
                  axis.text = element_text(size = 8),
                  plot.background = element_blank(),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  strip.text.x = element_blank(),
                  legend.position = 'bottom')+
            scale_fill_brewer(palette='Greys') + 
                labs(y = 'proportion', x = 'dominant cycle')+
            facet_wrap(~stage, nrow = 1, ncol = 2)

# arrange figure 
fig1 <- plot_grid(indi_plot, sp_plot,ncol = 1, nrow = 2)

fig1

# save figure
ggsave(file = paste0(figpath, '/fig1.png'),
       plot = fig1,
       width = 5, 
       height = 4, 
       units = 'in')
```
Sangasanga: sub-annual flowering and fruiting cycles were most frequently observed across individuals and species\
Vatovavy: supra-annual flowering and fruiting cycles were most frequently observed across individuals, sub-annual and no dominance in cycles was the most frequently observed flowering cycles across species and no dominance in cycles was the most frequently observed fruiting cycles across species.

Chi square and Fisher exact test to examine whether plants in these communities different distributions of cycle types 
```{r, echo = T, warning = F, message = F}
# Run Chi Squared statistiscs to see if we observe unequal amount of dominant cycles 

# run for Vatovavy, flowering individual 
chisq.test(dompop_summary$n[which(dompop_summary$dataset == 'Vatovavy' & 
                                    dompop_summary$stage == 'flower' &
                                    dompop_summary$scale=='individual')])

# run for Sangasanga, flowering individual 
chisq.test(dompop_summary$n[which(dompop_summary$dataset == 'Sangasanga' & 
                                    dompop_summary$stage == 'fruit' &
                                    dompop_summary$scale=='individual')])

# run for Vatovavy, fruiting individual 
chisq.test(dompop_summary$n[which(dompop_summary$dataset == 'Vatovavy' & 
                                    dompop_summary$stage == 'fruit' &
                                    dompop_summary$scale=='individual')])

# run for Sangasanga, fruiting individual 
chisq.test(dompop_summary$n[which(dompop_summary$dataset == 'Sangasanga' & 
                                    dompop_summary$stage == 'fruit' &
                                    dompop_summary$scale=='individual')])


# run for Vatovavy, flowering species
fisher.test(cbind(dompop_summary$n[which(dompop_summary$dataset == 'Vatovavy' & 
                                    dompop_summary$stage == 'flower' &
                                    dompop_summary$scale=='species')], rep(2,5)))

# run for Sangasanga, fruiting species
fisher.test(cbind(dompop_summary$n[which(dompop_summary$dataset == 'Vatovavy' & 
                                    dompop_summary$stage == 'fruit' &
                                    dompop_summary$scale=='species')], rep(2,5)))

# run for Sangasanga, flowering species
fisher.test(cbind(dompop_summary$n[which(dompop_summary$dataset == 'Sangasanga' & 
                                    dompop_summary$stage == 'flower' &
                                    dompop_summary$scale=='species')], rep(2,5)))

# run for Sangasanga, fruiting species
fisher.test(cbind(dompop_summary$n[which(dompop_summary$dataset == 'Sangasanga' & 
                                    dompop_summary$stage == 'fruit' &
                                    dompop_summary$scale=='species')], rep(2,5)))
```

Run post-hoc binomial test to see if dominant cycles differed from other cycles 
```{r, echo = T, warning = F, message = F}
# function for identifying the dominant category and conducting binomial test
dom_binom <- function(df, d, st, sc){
  # df: dominant cycle table 
  # d: specify which forest: 'Sangasanga' or 'Vatovavy'
  # st: specify which stage: 'flower' or 'fruit'
  # sc: specify which scale: 'individual' or 'species'
  df <- df %>% filter(dataset == d & stage == st & scale == sc) # filter dataframe by forest, stage and scale
  dom_cycle <- df$cycle_category[which(df$n == max(df$n))][1] # get the dominant cycle category
  df$dn <- df$n[which(df$cycle_category == dom_cycle)] # get the number of times that cycle category occured
  df$dom_cycle <- dom_cycle # add a column that specifies the dominant cycle category
  # get rid of rows in which the cycle category is equal to the dominant category
  df <- df[-which(df$cycle_category == dom_cycle),] 
  df$dtot <- df$n + df$dn # sum up the number of times the plants exhibited nondomimant cycles collectively
  df$sig <- sapply(1:nrow(df),
                   function(x) binom.test(x=df$dn[x],
                                          n=df$dtot[x],
                                          p = 0.05, 
                                          alternative='greater')$p.value) # run the binomial test
  return(df)
}

# create empty table to save results
binom_results <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(binom_results) <- c('dn','dom_cycle','dtot','sig') # assign column names

# run binomial test for each combination of scale, stage and forest
for (i in c('individual','species')) {
  for (j in c('flower','fruit')) {
    for (k in c('Sangasanga','Vatovavy')) {
      df <- dom_binom(dompop_summary,k,j,i)
      binom_results <- rbind(binom_results,df)
    }
  }
}

print(binom_results) # output  results 

# save results
write.csv(binom_results, 'dom_cycle_binomial_test.csv')
```

Visualize the distribution of median cycle length, duration and interval for individuals and species
```{r, echo = T, warning = F, message = F}
# summary the median cycle length, duration and interval per species
community_sum <- cycle_summary %>% group_by(scale, stage, dataset) %>% 
                                  summarise(n = n(), length_mean = mean(length_median,na.rm=T),
                                            length_sd = sd(length_median,na.rm=T),
                                            length_mad = mad(length_median,na.rm=T), 
                                            length_median = median(length_median,na.rm=T), 
                                            duration_mean = mean(duration_median,na.rm=T),
                                            duration_sd = sd(duration_median,na.rm=T),
                                            duration_mad = mad(duration_median,na.rm=T),
                                            duration_median = median(duration_median,na.rm=T), 
                                            interval_mean = mean(interval_median,na.rm=T),
                                            interval_sd = sd(interval_median,na.rm=T),
                                            interval_mad = mad(interval_median,na.rm=T),
                                            interval_median = median(interval_median,na.rm=T))
                                            
community_sum

# plot for Sangasanga cycle lengths
length_plot <- ggplot(cycle_summary %>% filter(dataset == 'sssw'), 
                      aes(x = length_median)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=length_median, linetype = 'dashed')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=length_median - length_mad, linetype = 'dotted')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=length_median + length_mad, linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'length (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, nrow = 4, ncol = 1, strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), scales = 'free')

# plot for Sangasanga cycle duration
dur_plot <- ggplot(cycle_summary  %>% filter(dataset == 'sssw'), 
                   aes(x = duration_median)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=duration_median, linetype = 'dashed')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=duration_median - duration_mad, linetype = 'dotted')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=duration_median + duration_mad, linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'duration (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, nrow = 4, ncol = 1, strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), scales = 'free')
  
# plot for Sangasanga cycle interval
int_plot <- ggplot(cycle_summary  %>% filter(dataset == 'sssw'), 
                   aes(x = interval_median)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=interval_median, linetype = 'dashed')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=interval_median - interval_mad, linetype = 'dotted')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'sssw'), 
                             aes(xintercept=interval_median + interval_mad, linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'interval (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, nrow = 4, ncol = 1, strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), scales = 'free')

# arrange figure 
cycle_flower <- grid.arrange(length_plot, dur_plot, int_plot, ncol = 3, nrow = 1)

# save plot
ggsave(file = paste0(figpath, '/FigS1.png'),
       plot = cycle_flower,
       width = 6.5, 
       height = 4, 
       units = 'in')

# output figure 
cycle_flower
```

```{r, echo = T, warning = F, message = F}
# plot for Vatovavy cycle lengths
length_plot <- ggplot(cycle_summary %>% filter(dataset == 'vvsw'), 
                      aes(x = length_median)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=length_median, 
                                 linetype = 'dashed')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=length_median - length_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=length_median + length_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'length (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, 
                             nrow = 4, 
                             ncol = 1, 
                             strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), 
                             scales = 'free')

# plot for Vatovavy cycle duration
dur_plot <- ggplot(cycle_summary  %>% filter(dataset == 'vvsw'), 
                   aes(x = duration_median)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=duration_median, 
                                 linetype = 'dashed')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=duration_median - duration_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=duration_median + duration_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'duration (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, nrow = 4, ncol = 1, strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), scales = 'free')
  
# plot for Vatovavy cycle interval
int_plot <- ggplot(cycle_summary  %>% filter(dataset == 'vvsw'), aes(x = interval_median)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=interval_median, 
                                 linetype = 'dashed')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=interval_median - interval_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sum %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=interval_median + interval_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'interval (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, 
                             nrow = 4, 
                             ncol = 1, strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), 
                             scales = 'free')

# arrange figure 
cycle_flower <- grid.arrange(length_plot, dur_plot, int_plot, ncol = 3, nrow = 1)

# save plot
ggsave(file = paste0(figpath, '/FigS2.png'),
       plot = cycle_flower,
       width = 6.5, 
       height = 4, 
       units = 'in')

# output figure 
cycle_flower
```

Compare cycle characteristics between communities
```{r, echo=T, warning=F, message=F}
community_diff <- community_sum[c(2,4,6,8),5:ncol(community_sum)] - community_sum[c(1,3,5,7),5:ncol(community_sum)]
community_diff$scale <- c('individual','individual','species','species')
community_diff$stage <- c('flower','fruit','flower','fruit')

# function to generate test statistic to compare difference in cycle characteristics between communities
perm_cycle_char <- function(df,sc,st){
  # df = dataframe
  # sc = individual or species
  # st = flower or fruit
  subdf <- cycle_summary %>% filter(sc == scale & st == stage) # subset the scale and stage 
  subdf$dataset <- sample(subdf$dataset, length(subdf$dataset),replace = F) # permute the community labels 
  # recaluate the the cycle characteristics
  submetrics <- subdf %>% group_by(dataset) %>% summarise(length_mean = mean(length_median,na.rm=T),
                                            length_sd = sd(length_median,na.rm=T),
                                            length_mad = mad(length_median,na.rm=T), 
                                            length_median = median(length_median,na.rm=T), 
                                            duration_mean = mean(duration_median,na.rm=T),
                                            duration_sd = sd(duration_median,na.rm=T),
                                            duration_mad = mad(duration_median,na.rm=T),
                                            duration_median = median(duration_median,na.rm=T), 
                                            interval_mean = mean(interval_median,na.rm=T),
                                            interval_sd = sd(interval_median,na.rm=T),
                                            interval_mad = mad(interval_median,na.rm=T),
                                            interval_median = median(interval_median,na.rm=T))
  # generate test statistic by taking the difference between the Vatovavy and Sangasanga cycle characteristics
  commdiff <- submetrics[2,2:ncol(submetrics)] - submetrics[1,2:ncol(submetrics)] 
  return(commdiff) # return the result
}

# function to get p-value
perm_pvalue <- function(obs_df, perm_df,sc, st, metric){
  subdf <- obs_df %>% filter(stage == st & scale == sc)
  metname <- colnames(obs_df)[metric]
  ref <- subdf[1,metric]
  # find the p-value
  perm_metric <- perm_df[,metname] # extract permuted test stat
  pval <- sum(abs(perm_metric) >= abs(ref))/length # get proportion of values that are more extreme than the absolute value of the observed metric, since a two tailed test is performed
  pval_df <- data.frame(scale = sc, stage = st,metric = metname, obs_value = ref, pvalue = pval) # output info
  return(pval_df)
}

# calculate for flower individual
fli_cycle_comp <- foreach(i = 1:9999, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  comm_diff <- perm_cycle_char(cycle_summary, 'individual','flower')
  return(comm_diff)
}

# add stage and scale info
fli_cycle_comp$scale <- 'individual'
fli_cycle_comp$stage <- 'flower'

# add observed value
fli_cycle_comp <- rbind(fli_cycle_comp,community_diff[which(community_diff$stage =='flower' & community_diff$scale == 'individual'),])

write.csv(fli_cycle_comp,'permutations_flower_individual_characteristics.csv') # save permutations

fli_pvalues <- foreach(i = 1:12, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  pmetrics <- perm_pvalue(community_diff, fli_cycle_comp, 'individual','flower',i)
  return(pmetrics)
}

# calculate for fruit individual
fri_cycle_comp <- foreach(i = 1:9999, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  comm_diff <- perm_cycle_char(cycle_summary, 'individual','fruit')
  return(comm_diff)
}

# add scale and stage info
fri_cycle_comp$scale <- 'individual'
fri_cycle_comp$stage <- 'fruit'

# add observed value
fri_cycle_comp <- rbind(fri_cycle_comp,community_diff[which(community_diff$stage =='fruit' & community_diff$scale == 'individual'),])

# save output
write.csv(fri_cycle_comp,'permutations_fruit_individual_characteristics.csv')

fri_pvalues <- foreach(i = 1:12, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  pmetrics <- perm_pvalue(community_diff, fri_cycle_comp, 'individual','fruit',i)
  return(pmetrics)
}


#calculate for flower species
flsp_cycle_comp <- foreach(i = 1:9999, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  comm_diff <- perm_cycle_char(cycle_summary, 'species','flower')
  return(comm_diff)
}

# add stage and scale info
flsp_cycle_comp$scale <- 'species'
flsp_cycle_comp$stage <- 'flower'

# add observed value
flsp_cycle_comp <- rbind(flsp_cycle_comp,community_diff[which(community_diff$stage =='flower' & community_diff$scale == 'species'),])

write.csv(flsp_cycle_comp,'permutations_flower_species_characteristics.csv')

flsp_pvalues <- foreach(i = 1:12, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  pmetrics <- perm_pvalue(community_diff, flsp_cycle_comp, 'species','flower',i)
  return(pmetrics)
}

# calculate for fruit species
frsp_cycle_comp <- foreach(i = 1:9999, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  comm_diff <- perm_cycle_char(cycle_summary, 'species','fruit')
  return(comm_diff)
}

# add scale and stage info
frsp_cycle_comp$scale <- 'species'
frsp_cycle_comp$stage <- 'fruit'

# add observed value
frsp_cycle_comp <- rbind(frsp_cycle_comp,community_diff[which(community_diff$stage =='fruit' & community_diff$scale == 'species'),])

write.csv(frsp_cycle_comp,'permutations_fruit_species_characteristics.csv')

frsp_pvalues <- foreach(i = 1:12, .combine = 'rbind', .packages = 'dplyr') %dopar%{
  pmetrics <- perm_pvalue(community_diff, frsp_cycle_comp, 'species','fruit',i)
  return(pmetrics)
}

comm_char_pvalues <- do.call("rbind", list(fli_pvalues,fri_pvalues,flsp_pvalues,frsp_pvalues))
write.csv(comm_char_pvalues, 'permutation_test_pvalue_cycle_characteristic_comparison.csv')
```

Compare median cycle length, duration and interval between communities
```{r, echo = T, warning = F, message = F}
# create empty table to save results
forest_comp <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(forest_comp) <- c('metric','stage','scale','tstate','dof','pval') # assign column names

# for each scale, stage and metric
for (i in c('individual','species')){
  for (j in c('flower','fruit')){
    for (k in c('length_median','duration_median','interval_median')){
      subdf <- cycle_summary %>% filter(scale == i & stage == j) # subset the stage and scale
      tdf <- t.test(subdf[which(subdf$dataset =='sssw'),k],subdf[which(subdf$dataset =='vvsw'),k]) # run t-test
      # add t-test output stats to data table 
      forest_comp[nrow(forest_comp)+1,] <- c(k,j,i,tdf$statistic, tdf$parameter, tdf$p.value) 
    }
  }
}

# save the t-test results
write.csv(forest_comp,'forest_comp_comm_phen.csv')
```

**Which individuals and species had the shortest and longest cycle length, duration and interval?**
```{r, echo = T, warning = F, message = F}
# longest individual flowering cycle
(cycle_summary %>% filter(scale =='individual' & stage == 'flower') %>% 
                   filter(length_max == max(length_max)))[,c('Species','Latin','length_max')]

# longest individual flowering duration
(cycle_summary %>% filter(scale =='individual' & stage == 'flower') %>% 
                   filter(duration_max == max(duration_max)))[,c('Species','Latin','duration_max')]

# longest individual flowering interval
(cycle_summary %>% filter(scale =='individual' & stage == 'flower') %>% 
                   filter(interval_max == max(interval_max)))[,c('Species','Latin', 'interval_max')]

# longest species flowering cycle
(cycle_summary %>% filter(scale =='species' & stage == 'flower') %>% 
                   filter(length_max == max(length_max)))[,c('Species','Latin','length_max')]

# longest species flowering duration
(cycle_summary %>% filter(scale =='species' & stage == 'flower') %>% 
                   filter(duration_max == max(duration_max)))[,c('Species','Latin','duration_max')]

# longest species flowering interval
(cycle_summary %>% filter(scale =='species' & stage == 'flower') %>% 
                   filter(interval_max == max(interval_max)))[,c('Species','Latin', 'interval_max')]

# longest individual fruiting cycle
(cycle_summary %>% filter(scale =='individual' & stage == 'fruit') %>% 
                   filter(length_max == max(length_max)))[,c('Species','Latin','length_max')]

# longest individual fruiting duration
(cycle_summary %>% filter(scale =='individual' & stage == 'fruit') %>% 
                   filter(duration_max == max(duration_max)))[,c('Species','Latin','duration_max')]

# longest individual fruiting interval
(cycle_summary %>% filter(scale =='individual' & stage == 'fruit') %>% 
                   filter(interval_max == max(interval_max)))[,c('Species','Latin', 'interval_max')]

# longest species fruiting cycle
(cycle_summary %>% filter(scale =='species' & stage == 'fruit') %>% 
                   filter(length_max == max(length_max)))[,c('Species','Latin','length_max')]

# longest species fruiting duration
(cycle_summary %>% filter(scale =='species' & stage == 'fruit') %>% 
                   filter(duration_max == max(duration_max)))[,c('Species','Latin','duration_max')]

# longest species fruiting interval
(cycle_summary %>% filter(scale =='species' & stage == 'fruit') %>% 
                   filter(interval_max == max(interval_max)))[,c('Species','Latin', 'interval_max')]
```

Visualize and and get summary statistics for cycle regularity
```{r, echo = T, warning = F, message = F}
# regularity is measured as the variation in cycle length, duration and interval 
# summary the standard deviation of the cycle length, duration and interval 
community_sd <- cycle_summary %>% group_by(scale, stage, dataset) %>% 
                                  summarise(length_min = min(length_sd,na.rm=T),
                                            length_max = max(length_sd,na.rm=T),
                                            length_mean = mean(length_sd,na.rm=T),
                                            length_mad = mad(length_sd,na.rm=T), 
                                            length_median = median(length_sd,na.rm=T),
                                            length_sd = sd(length_sd,na.rm=T),
                                            duration_min = min(duration_sd,na.rm=T),
                                            duration_max = max(duration_sd,na.rm=T),
                                            duration_mean = mean(duration_sd,na.rm=T),
                                            duration_mad = mad(duration_sd,na.rm=T),
                                            duration_median = median(duration_sd,na.rm=T), 
                                            duration_sd = sd(duration_sd,na.rm=T),
                                            interval_min = min(interval_sd,na.rm=T),
                                            interval_max = max(interval_sd,na.rm=T),
                                            interval_mean = mean(interval_sd,na.rm=T),
                                            interval_mad = mad(interval_sd,na.rm=T),
                                            interval_median = median(interval_sd,na.rm=T),
                                            interval_sd = sd(interval_sd,na.rm=T))
                                            
# plot the standard deviation in the cycle length for Sangasanga
length_plot <- ggplot(cycle_summary %>% filter(dataset == 'sssw'), 
                      aes(x = length_sd)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=length_median, 
                                 linetype = 'dashed')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=length_median - length_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=length_median + length_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'length (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, nrow = 4, ncol = 1, strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), scales = 'free')

# plot the standard deviation in the cycle duration for Sangasanga
dur_plot <- ggplot(cycle_summary %>% filter(dataset == 'sssw'), 
                   aes(x = duration_sd)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=duration_median, 
                                 linetype = 'dashed')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=duration_median - duration_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=duration_median + duration_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'duration (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, 
                             nrow = 4, 
                             ncol = 1, 
                             strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), 
                             scales = 'free')
  
# plot the standard deviation in the cycle interval for Sangasanga
int_plot <- ggplot(cycle_summary %>% filter(dataset == 'sssw'), 
                   aes(x = interval_sd)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=interval_median, 
                                 linetype = 'dashed')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=interval_median - interval_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'sssw'), 
                             aes(xintercept=interval_median + interval_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'interval (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, 
                             nrow = 4, 
                             ncol = 1, 
                             strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), 
                             scales = 'free')

# arrange figure 
cycle_flower <- grid.arrange(length_plot, dur_plot, int_plot, ncol = 3, nrow = 1)

# save plot
ggsave(file = paste0(figpath, '/FigS3.png'),
       plot = cycle_flower,
       width = 6.5, 
       height = 4, 
       units = 'in')

# output figure 
cycle_flower
```

```{r, echo = T, warning = F, message = F}
# plot the standard deviation in the cycle length for Vatovavy
length_plot <- ggplot(cycle_summary %>% filter(dataset == 'vvsw'), 
                      aes(x = length_sd)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=length_median, linetype = 'dashed')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=length_median - length_mad, linetype = 'dotted')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=length_median + length_mad, linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'length (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, 
                             nrow = 4, 
                             ncol = 1, 
                             strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), 
                             scales = 'free')

# plot the standard deviation in the cycle duration for Vatovavy
dur_plot <- ggplot(cycle_summary %>% filter(dataset == 'vvsw'), 
                   aes(x = duration_sd)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=duration_median, 
                                 linetype = 'dashed')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=duration_median - duration_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=duration_median + duration_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'duration (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, 
                             nrow = 4, 
                             ncol = 1, 
                             strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), 
                             scales = 'free')
  
# plot the standard deviation in the cycle interval for Vatovavy
int_plot <- ggplot(cycle_summary %>% filter(dataset == 'vvsw'), 
                   aes(x = interval_sd)) +
                  geom_histogram(color = 'black', fill = 'darkgray') +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=interval_median, linetype = 'dashed')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=interval_median - interval_mad, 
                                 linetype = 'dotted')) +
                  geom_vline(data=community_sd %>% filter(dataset == 'vvsw'), 
                             aes(xintercept=interval_median + interval_mad, 
                                 linetype = 'dotted')) +
                  theme_bw() +
                  theme(legend.position = 'none', 
                        text = element_text(size = 8), 
                        axis.text = element_text(size = 8)) +
                  labs(x = 'interval (months)', y = 'frequency') +
                  facet_wrap(~scale + stage, 
                             nrow = 4, 
                             ncol = 1, 
                             strip.position = 'right', 
                             labeller = label_wrap_gen(multi_line=FALSE), 
                             scales = 'free')

# arrange figure 
cycle_flower <- grid.arrange(length_plot, dur_plot, int_plot, ncol = 3, nrow = 1)

# save figure 
ggsave(file = paste0(figpath, '/FigS4.png'),
       plot = cycle_flower,
       width = 6.5, 
       height = 4, 
       units = 'in')

# output figure
cycle_flower

# output table 
community_sd
```

Compare median cycle length, duration and interval between communities
```{r, echo = T, warning = F, message = F}
# create empty table to save results
forest_comp <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(forest_comp) <- c('metric','stage','scale','tstate','dof','pval') # assign column names

# for each scale, stage and metric
for (i in c('individual','species')){
  for (j in c('flower','fruit')){
    for (k in c('length_median','duration_median','interval_median')){
      subdf <- cycle_summary %>% filter(scale == i & stage == j) # subset the stage and scale
      tdf <- t.test(subdf[which(subdf$dataset =='sssw'),k],subdf[which(subdf$dataset =='vvsw'),k]) # run t-test
      # add t-test output stats to data table 
      forest_comp[nrow(forest_comp)+1,] <- c(k,j,i,tdf$statistic, tdf$parameter, tdf$p.value) 
    }
  }
}

# save the t-test results
write.csv(forest_comp,'forest_comp_comm_phen_median.csv')
```

Compare cycle length, duration and interval regularity between communities
```{r, echo = T, warning = F, message = F}
# create empty table to save results
forest_comp <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(forest_comp) <- c('metric','stage','scale','tstate','dof','pval') # assign column names

# for each scale, stage and metric
for (i in c('individual','species')){
  for (j in c('flower','fruit')){
    for (k in c('length_sd','duration_sd','interval_sd')){
      subdf <- cycle_summary %>% filter(scale == i & stage == j)  # subset the stage and scale
      tdf <- t.test(subdf[which(subdf$dataset =='sssw'),k],subdf[which(subdf$dataset =='vvsw'),k]) # run t-test
      forest_comp[nrow(forest_comp)+1,] <- c(k,j,i,tdf$statistic, tdf$parameter, tdf$p.value)
    }
  }
}

write.csv(forest_comp,'forest_comp_comm_phen_sd.csv')
```
